---
title: "W203 Lab 2 Research Proposal"
author: "Ed Brown, Daphne Lin, Linh Tran, Lisa Wu"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
 \usepackage{wrapfig}
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages and set options, include=FALSE}
library(tidyverse) 
library(stargazer)
library(sandwich)
library(lmtest)

theme_set(theme_bw())
```


## Research Question

Violent crime has been increasing in the Unitied States since reaching a low in 2014 \footnote{https://crime-data-explorer.app.cloud.gov/pages/explorer/crime/crime-trend}. The Principal Components Consulting Group has been contracted by the U.S. Department of Health and Human Services (HHS) as part of their "Healthy People 2030 Program" \footnote{https://health.gov/healthypeople/priority-areas/social-determinants-health/literature-summaries/crime-and-violence} to identify primary causes of violent crime which will inform future intervention from (HHS) in drafting proposal and policies to curtail the rising crime reports. Specifically, we are examining incidence of violent crime per population and whether family stability affects (reduces) the incidence of violent crime.   

## Data Source

Our data source is the "Communities and Crime" data set located at the UCI Machine Learning Repository \footnote{(Dua, D. and Graff, C. (2019). UCI Machine Learning Repository [http://archive.ics.uci.edu/ml]. Irvine, CA: University of California, School of Information and Computer Science.)}. We have chosen the outcome variable of ViolentCrimesPerPop, which is defined as the total number of violent crimes per 100K population. We are currently evaluating several input variables to operationalize the concept of family stability.  Based on our background research, preliminary exploratory data analysis and statistical testing, we have chosen two variables - PctIlleg (Percent Illegitimate Children), and TotalPctDiv (Total Divorce Percent) to represent the family status. High PctIlleg and high TotalPctDiv represent low (or lack of) family stability, and low PctIlleg and low TotalPctDiv represent more family stability.
 
## Unit of Observation

Each row represents cross sectional data from counties in the United States. There is one observation for each county. We have data from 1994 of 3006 counties in the United States in the raw data set before any filtering. Based on the description provided by the data source, all numeric data was normalized into the decimal range 0.00-1.00 using an Unsupervised, equal-interval binning method. Attributes retain their distribution and skew (hence for example the population attribute has a mean value of 0.06 because most communities are small). E.g. An attribute described as 'mean people per household' is actually the normalized (0-1) version of that value\footnote{https://archive.ics.uci.edu/ml/datasets/communities+and+crime}.


