---
title: "W203 Lab 2 Research Proposal"
author: "Ed Brown, Daphne Lin, Linh Tran, Lisa Wu"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
 \usepackage{wrapfig}
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages and set options, include=FALSE}
library(tidyverse) 
library(stargazer)
library(sandwich)
library(lmtest)

theme_set(theme_bw())
```

## Research Question

Principal Components Consulting Group has been contracted by a big three automotive company to identify key factors in developing a pricing model for new automobile introduction. Specifically, we are explaining the manufacturer suggested retail price (MSRP) in terms of engine displacement.    

Engine displacement is a design parameter for automobiles given the available engine space, automobile size and desired performance characteristics of the automobile. With smaller more economical cars having smaller engines and larger more luxurious or sporty vehicles having larger engines.  

## Data Source

Our data source is from Kaggle\footnote{https://www.kaggle.com/datasets/gagandeep16/car-sales}. We have chosen the outcome variable of price_in_thousands, which is the MSRP price for a new vehicle. We are currently evaluating several input variables to operationalize this test.  Based on our background research, preliminary exploratory data analysis and statistical testing, we have chosen engine displacement as the primary variable. We will augment this base case with additional categorical variables to refine the model.
 
## Unit of Observation

Each row of the data set represents cross sectional data and is a single vehicle model available for purchase. There is one observation per vehicle. There are a couple aggregate columns in the data e.g. resale, however, we will not leverage them as part of this study.



